Nock for Everyday Coders

### 2025-04-19 mss.tex
Post review copy.

How do you say which slot number you want from the tree?  We say that:

@@

This is saying to run \lstinline[style=inlinecode]{*[a b]}, and then make that the head of a new subject, with the old subject \lstinline[style=inlinecode]{a} as the new subject's tail.

@@

If ``grab a chunk of code from a subject and then run it against the subject" sounds a lot like getting an arm from a core in Hoon, that's because it is.

The ``new subjects" created by \lstinline[style=inlinecode]{*[a c]} are cores, and the things selected from memory slots by \lstinline[style=inlinecode]{b} are arms.

@@

\begin{enumerate}
  \item  Assign the mold gate \lstinline[style=inlinecode]{?} to the face \lstinline[style=inlinecode]{boolean-mold}.
  \item  Take a look at what's inside it...looks like the head is an arm...let's print out its source code!
  \item  Print out that source code by calling the head.
\end{enumerate}

So, we know the below code is a gate that evaluates its sample, and returns it if it's a boolean, and crashes otherwise. Let's see how that works.

@@

What is the subject and what is at mem slot \lstinline[style=inlinecode]{6}? This code is the arm of a gate, so the subject is that gate/core: \lstinline[style=inlinecode]{[battery [sample payload]]}. We are looking at the battery right now, and so \lstinline[style=inlinecode]{[@0@ 6]} yanks the head of the tail...the \lstinline[style=inlinecode]{sample}!

@@

Otherwise, we run the 2nd branch of the if-else:

@@

In this section, we've seen how to use all the remaining opcodes, which build Nock up to a slightly more expressive level. We also saw how Hoon cores start to arise pretty naturally out of the Nock primitives, especially \lstinline[style=inlinecode]{@8@} and \lstinline[style=inlinecode]{@9@}. Then we walked through real production Nock code to show that everything we've learned so far works exactly as expected in the wild.

In the next section, you'll learn how to write real programs in Nock, and compose those programs to make new ones.

Finally, hopefully you now see that, whatever its other limitations, Nock is not particularly obscurantist, and is fairly straightforward to parse, once you understand its syntax and idioms.

@@

At this point, the interpreter expands ``inside-out", starting from the furthest inside pseudocode operators. In this case, those are \lstinline[style=inlinecode]{/[1 50]} and \lstinline[style=inlinecode]{50}, so it does those:

@@

\begin{lstlisting}[style=listingcode]
.*([50 51] [11 369 @0@ 2])
\end{lstlisting}

@@

\section{The Core as Design Patterns}

@@

\begin{lstlisting}[style=listingcode]
*[subject 2 formula-to-set-up-core formula-against-the-core]
\end{lstlisting}

>>>>

\noindent{}
This permits the address to be defined within a particular subtree as well as within the overall tree (Figure \ref{fig:treefull}).

\begin{figure}[hbt!]
  \centering
  \caption{A binary tree with labeled node addresses to several layers. \label{fig:treefull}}
  \begin{tikzpicture}[
    ->, >=stealth,
    level 1/.style={sibling distance=8cm, level distance=1.5cm},
    level 2/.style={sibling distance=4cm, level distance=1.5cm},
    level 3/.style={sibling distance=2cm, level distance=1.5cm},
    level 4/.style={sibling distance=1cm, level distance=1.5cm},
    scale=0.6, transform shape,
    subtree/.style={circle, draw, fill=gray!20, minimum size=1cm},
    treenode/.style={circle, draw, fill=white, minimum size=1cm}
  ]
    \node [subtree] {\Large 1}
    child {
      node [subtree] {\Large 2}
      child {
        node [subtree] {\Large 4}
        child {
          node [subtree] {\Large 8}
          child { node [treenode] {\Large 16} }
          child { node [treenode] {\Large 17} }
        }
        child {
          node [subtree] {\Large 9}
          child { node [treenode] {\Large 18} }
          child { node [treenode] {\Large 19} }
        }
      }
      child {
        node [subtree] {\Large 5}
        child {
          node [subtree] {\Large 10}
          child { node [treenode] {\Large 20} }
          child { node [treenode] {\Large 21} }
        }
        child {
          node [subtree] {\Large 11}
          child { node [treenode] {\Large 22} }
          child { node [treenode] {\Large 23} }
        }
      }
    }
    child {
      node [subtree] {\Large 3}
      child {
        node [subtree] {\Large 6}
        child {
          node [subtree] {\Large 12}
          child { node [treenode] {\Large 24} }
          child { node [treenode] {\Large 25} }
        }
        child {
          node [subtree] {\Large 13}
          child { node [treenode] {\Large 26} }
          child { node [treenode] {\Large 27} } % Fixed duplicate 26 to 27
        }
      }
      child {
        node [subtree] {\Large 7}
        child {
          node [subtree] {\Large 14}
          child { node [treenode] {\Large 28} }
          child { node [treenode] {\Large 29} }
        }
        child {
          node [subtree] {\Large 15}
          child { node [treenode] {\Large 30} }
          child { node [treenode] {\Large 31} }
        }
      }
    };
  \end{tikzpicture}
\end{figure}

That is, how do you say which slot number you want from a given tree?  We say that:

@@

This says to run \lstinline[style=inlinecode]{*[a b]}, and then make that the head of a new subject, with the old subject \lstinline[style=inlinecode]{a} as the new subject's tail.

@@

If ``grab a chunk of code from a subject and then run it against the subject" sounds a lot like getting an arm from a core in Hoon, that's because it is.  Nock opcode \lstinline[style=inlinecode]{@9@} is hand-in-glove with Hoon's core/arm structure.  The ``new subjects" created by \lstinline[style=inlinecode]{*[a c]} are cores, and the things selected from memory slots by \lstinline[style=inlinecode]{b} are arms.

@@

\begin{enumerate}
  \item  Assign the mold gate denoted by \lstinline[style=inlinecode]{?} to the face \lstinline[style=inlinecode]{boolean-mold}.
  \item  Examine what's inside, seeing that the head is an arm.
  \item  Print out that source code by calling the head.
\end{enumerate}

\noindent{}
We know the below code is a gate that evaluates its sample, and returns it if it's a boolean, and crashes otherwise. Let's see how that works.

@@

What is the subject and what is at mem slot \lstinline[style=inlinecode]{6}? This code is the arm of a gate, so the subject is that gate/core: \lstinline[style=inlinecode]{[battery [sample payload]]}. We are looking at the battery right now, and so \lstinline[style=inlinecode]{[@0@ 6]} yanks the head of the tail, or the \lstinline[style=inlinecode]{sample}!

@@

Otherwise, we run the second branch of the \lstinline[style=inlinecode]{if}–\lstinline[style=inlinecode]{else} conditional:

@@

In this section, we've seen how to use almost all of the remaining opcodes, which build Nock up to a slightly more expressive level. We also saw how Hoon cores start to arise pretty naturally out of the Nock primitives, especially \lstinline[style=inlinecode]{@8@} and \lstinline[style=inlinecode]{@9@}. Then we walked through real production Nock code to show that everything we've learned so far works exactly as expected in the wild.  In the next section, you'll learn how to write real programs in Nock, and compose those programs to make new ones.  Finally, hopefully you now see that, whatever its other limitations, Nock is not particularly obscurantist, and is fairly straightforward to parse, once you understand its syntax and idioms.

@@

At this point, the interpreter expands ``inside-out", starting from the deepest operators. In this case, those are \lstinline[style=inlinecode]{/[1 50]} and \lstinline[style=inlinecode]{50}, so it does those:

@@

\begin{lstlisting}[style=listingcode]
.*([50 51] [@11@ 369 @0@ 2])
\end{lstlisting}

@@

\section{The Core as Design Pattern}

@@

\begin{lstlisting}[style=listingcode]
*[subject @2@ core-formula call-formula]
\end{lstlisting}
